Combinatorics of the geometry of Wilson loop diagrams II: Grassmann necklaces, dimensions, and denominators

### 2020-12-13 GWLDCI.tex
reviewed Karen  changes;  addressed comments

Matroid theory relates to the study of Wilson loop diagrams as follows. In \cite[section 3]{wilsonloop}, Agarwala and Marin-Amat show that every weakly admissible Wilson loop diagram $W$ with $k$ propagators defines a positroid $M(W)$ of rank $k$ on a base set of size $n$. \hlfix{Combined with Postnikov's work in {\cite{Postnikov}}, this shows that each admissible Wilson loop diagram corresponds to a convex subset of $\Gr(k,n)$}{fix citation. Ask Cameron?}, thus connecting the geometry of Wilson loop diagrams with the geometry underlying the Amplituhedron.

>>>>

Matroid theory relates to the study of Wilson loop diagrams as follows. In \cite[section 3]{wilsonloop}, Agarwala and Marin-Amat show that every weakly admissible Wilson loop diagram $W$ with $k$ propagators defines a positroid $M(W)$ of rank $k$ on a base set of size $n$. \hlfix{It is a direct consequence of Postnikov's work, {\cite[Theorem 6.5]{Postnikov}}, that positroid cells are convex subset of $\Gr(k,n)$}{I  hate  saying it this way, but I can't find it directly written anywhere}. Putting these together gives that the Wilson loop diagrams  correspond to convex subsets of $\Gr(k,n)$, connecting the geometry of Wilson loop diagrams with the geometry underlying the Amplituhedron.

### 2020-12-13 GWLDCII.tex
reviewed Karen  changes;  addressed comments

Meanwhile, a different body of work has studied SYM $N=4$ theory from the point of view of Feynman integrals in twistor space \cite{Adamo:2011pv,Adamo:2012xe,  Boels:2007qn, Bullimore:2010pj}. These integrals are calculated via holomorphic Wilson loops. As with the Amplituhedron, each Wilson loop represents the sum of a family of $N^kMHV$ Feynman diagrams, and adding up the corresponding sums of Feynman integrands associates a volume form to a convex space different from those in the decomposition the Amplituhedron \cite{CachazoSvrcekWitten, Hodges2013}.  \hlfix{The Wilson loop diagrams discussed in this paper are, loosely speaking, dual graphs of the $N^kMHV$ diagrams. While the Wilson loop diagrams contain the exact same information as the $N^kMHV$ diagrams for the purposes of understanding $N=4 SYM$ theory, this dual perspective is better suited to understanding the combinatorics and geometry associated to the theory.}{added to address request (3) of report M -- let's discuss, I don't think this makes sense as written} Therefore, \hlfix{we part with the notation in the physics literature, and name these dual graphs}{I'm not sure that's true} of $N^kMHV$ diagrams Wilson loop diagrams in order to emphasize the combinatorial and geometric significane of these object. In \cite{wilsonloop}, Agarwala and Marin-Amat uncovered a connection between the Feynman integrals developed in \hlfix{this}{What does this ``this'' refer to, also this sentence should probably start a new paragraph} literature and positroids, defined  as matroids that can be realized as an element of $\Gr(k,n)$. Other work has shown a connection between the volume filled by Wilson loop diagrams and the Amplituhedron \cite{casestudy, Amplituhedronsquared}. Specifically, Marcott shows that the space parametrized by the Wilson loop diagrams is  exactly the correct positroid cell in $\Gr(k,n)$, up to a set of measure zero \cite{WLDdim}.

In the above work and elsewhere, it is clear that while the space \hlfix{filled by}{``filled by'' vs ``parameterized by'' in the previous paragraph} the Wilson loop diagrams is related to the Amplituhedron, they are not the same space. For instance, it is conjectured that the Amplituhedron is orientable, whereas Agarwala and Marcott show in \cite{non-orientable} that the space parametrized by the Wilson loop diagrams as a subspace of $\Gr(k, n+1)$ can be seen as a non-orientable vector bundle over some submanifold of $\Gr(k,n)$. Furthermore, the cells associated to the Wilson loop diagrams are different from the cells associated to BCFW diagrams, which can be seen by comparing the results of \cite{AmplituhedronDecomposition} and \cite{casestudy}. This is an intriguing observation, as it is conjectured that the space parametrized by the Wilson loop diagrams will project down to something close to the Amplituhedron \cite{wilsonloop, Amplituhedronsquared}.

@@

This paper continues this body of work by establishing concrete connections between the combinatorics of the Wilson loop diagram as determined by the physics, and the combinatorics of the associated positroid cell as identified in \cite{wilsonloop}. In this paper, we give an algorithm to read the Grassmann necklace, one of the standard combinatorial characterizations of the positroid cell, directly off the Wilson loop diagram. We also show that the positroid cells associated to Wilson loop diagrams are all $3k$-dimensional, where $k$ is the number of propagators in the diagram; by contrast, the BCFW cells are all $4k$-dimensional \cite{Arkani-Hamed:2013jha}. \hlfix{This apparent contradiction}{I think we should say more in view of reviewer Y's comment 3 and we can probably phrase it more positively too} is resolved by the non-orientability result of \cite{non-orientable}.

@@

Finally, in Section~\ref{sec poles} we characterize the denominator of the integrand of a Wilson loop diagram in terms of its associated matrix and Grassmann necklace. We show that given the \hlfix{physically correct}{This was here before but can you remind me what it is hinting at?} representation of the positroid cell associated to a Wilson loop diagram, the irreducible factors of the denominator are exactly the set of irreducible factors of the minors of said representation given by the Grassmann necklace. This characterization is given explicitly by Algorithm~\ref{alg WLD to denom via GN} and Theorem~\ref{thm denom}.

@@

All three of these conditions arise from the physical interpretation of Wilson loop diagrams.  The first condition comes from the fact that these diagrams represent $N^k$MHV interctions, where $k +4 <n$, see for instance the explanation in the introduction of \cite{adamo:2011pr}.  The second condition can be derived from the integrals associated to the diagrams, \hlfix{as in.}{need end of sentence.}  Note that the second condition prohibits propagators that connect two adjacent edges. In our definition, $\cP$ is a set so we may not have two propagators starting and ending on the same pair of edges, but we have lost nothing of physical value by not allowing multiple propagators ending on the same pair of edges since these would also fail the second condition.   \cite[Theorem 1.13]{wilsonloop}\hlfix{.}{This cite was sitting as a sentence on its own and I don't know what it was for.}  The third condition comes from the fact that these Wilson loop diagrams are a representation of planar Feynman diagrams of $N=4$ SYM theory. For a description of the relationship between planar Feynman diagram and planar Wilson loops, see the discussion in \cite[section 3]{LipsteinMason}.
\hlfix{}{I changed the previous 2 paragraphs quite a bit, to separate the math intuition from the physical reasoning.}

@@

One may also interpret the functional $I(W)$ as associating a volume to each cell $\Sigma(W)$ \cite{wilsonloop, Amplituhedronsquared, HeslopStewart}, where the volume is determined by the external particles involved. Specifically, the integrand is written in terms of the elements of $C(W)$ as  \begin{equation} \label{eq:omega(W) def}\Omega(W) = \frac{\prod_{r=1}^{|\cP|} \prod_{v \in V(p_r)} \textrm{d}x_{p_r}}{R(W)} \;, \end{equation} multiplied by a $\delta$ function that evaluates the matrix $C(W)$ on the external particles. The denominator $R(W)$ is given in Definition~\ref{def R(W)} below. For the purposes of this paper, we restrict our attention to $\Omega(W)$ only, and ignore the $\delta$ function in the integrand of $I(W)$. The interested reader is referred to \cite{Adamo:2012xe,HeslopStewart,LipsteinMason} for more information about the integrand of $I(W)$. \sanote{not putting sections here, but I'd like a second thought -- what is this in response to?}

For a fixed number of propagators and particles, the tree level amplitude of the Wilson loop diagrams is computed as the sum of the $I(W)$ for all $W$ of the correct size. By {\cite{Adamo:2011pv,Adamo:2012xe,Arkani-Hamed:2013jha}} this sum is finite for all sets of external particles,\sanote{fix this citation} i.e.
\bas \mathcal{A}_{k, n}^{tree} = \sum_{\substack{W = (\cP, [n]), \\ |\cP| = k}} I(W) <\infty. \eas

This would certainly hold if  the singularities in the integrands $\Omega(W)$ cancel out in the sum; in particular, all poles arising from $R(W)$ must cancel. Specifically, it is conjectured that these poles cancel on the boundaries of the positroid cells $\Sigma(W)$, analogous to the analysis in \cite{Arkani-Hamed:2013jha}. This has been checked for particular classes of examples (e.g. \cite{casestudy, Amplituhedronsquared, HeslopStewart}) but has not been proven in general.\hlfix{}{Is this updated for the spurious vs physical poles?}

@@

Note that the positroid cells associated to the Amplituhedron are of dimension $4|\cP|$. In \cite[section 2]{non-orientable}, Agarwala and Marcott associate a Deodhar component\footnote{Since Deodhar components play no role in this paper, the interested reader is referred to \cite{non-orientable} for more details on this.} to each Wilson loop diagram, and show that the dimension of this Deodhar component is the sum of $|\cP|$ and the dimension of the associated positroid cell.
%\[|\cP| + \text{ the dimension of associated positroid cell}.\]
Combining this with Theorem~\ref{thm dim}, we see that the Deodhar component associated to a Wilson loop diagram is $4|\cP|$-dimensional, showing that the geometry underlying the Wilson loop diagrams is consistent with the geometry underlying the Amplituhedron.

Marcott in \cite[Theorem 8.4]{WLDdim} gives a different proof of the $3|\cP|$-dimensionality of $M(W)$ which is geometric and more elegant, but it is not easy to track the effect of a particular propagator.  It is worth noting that while the Amplituhedon is also defined by a projection of a set of $4|\cP|$ dimensional positroid cells, there is no evidence to suggest that the $3|\cP|$ dimensional cells of $\Gr(|\cP|, n)$ defiining the Wilson loop diagrams are related to the $4|\cP|$ dimensional cells defining the Amplituhedron.  \sanote{This para adresses point 3 of report Y. Should there be a similar paragraph in the intro? Should it be moved?}

\hlfix{Our approach is}{The transition between these paragraphs has been messed up a bit} much more explicit. By combining Algorithms~\ref{alg:put GN on WLD} and \ref{alg:GN to Le} (converting from Wilson loop diagram to Grassmann necklace, and Grassmann necklace to Le diagram respectively) we explicitly describe the effect of adding another propagator to a Wilson loop diagram in terms of the plusses of the associated Le diagrams. Recall that the dimension of a positroid cell is equal to the number of plusses in its associated Le diagram \cite[Theorem 6.5]{Postnikov}.  Thus understanding the effect of adding another propagator is sufficient to give a recursive proof of the $3|\cP|$-dimensionality of the cells.

@@

Since SYM $N=4$ is a finite theory {\cite{Alday:2007hr, ParkeTaylorids}}, the scattering amplitudes (and thus the holomorphic Wilson loops $\,\mathfrak{W}_{k,n}$) are finite.\sanote{Fix citations} That is, for a fixed set of external particles $\cZ$ we must have
\bas \sum_{W \text{ s.t. } |\cP| = k} I(W)(\cZ) <\infty \;.\eas
This holds if all poles of the integrands cancel. As a step towards this, in this section we prove some results about the structure of the denominator $R(W)$ of the integrand, as defined in Definition~\ref{def R(W)}.

>>>>

Meanwhile, a different body of work has studied SYM $N=4$ theory from the point of view of Feynman integrals in twistor space \cite{Adamo:2011pv,Adamo:2012xe,  Boels:2007qn, Bullimore:2010pj}. These integrals are calculated via holomorphic Wilson loops. As with the Amplituhedron, each Wilson loop represents the sum of a family of $N^kMHV$ Feynman diagrams, and adding up the corresponding sums of Feynman integrands associates a volume form to a convex space different from those in the decomposition the Amplituhedron \cite{CachazoSvrcekWitten, Hodges2013}.  \hlfix{The Wilson loop diagrams discussed in this paper are, loosely speaking, dual graphs of the $N^kMHV$ diagrams. While the Wilson loop diagrams contain the exact same information as the $N^kMHV$ diagrams for the purposes of understanding $N=4 SYM$ theory, this dual perspective is better suited to understanding the combinatorics and geometry associated to the theory.}{added to address request (3) of report M -- let's discuss, I don't think this makes sense as written} Therefore, \hlfix{we part with the notation in the physics literature, and name these dual graphs}{I'm not sure that's true} of $N^kMHV$ diagrams Wilson loop diagrams in order to emphasize the combinatorial and geometric significane of these object.

In \cite{wilsonloop}, Agarwala and Marin-Amat uncovered a connection between the Feynman integrals developed in the $N^kMHV$ \st{this}\todo{What does this ``this'' refer to, also this sentence should probably start a new paragraph} literature and positroids, defined  as matroids that can be realized as an element of $\Gr(k,n)$. Other work has shown a connection between the volume filled by Wilson loop diagrams and the Amplituhedron \cite{casestudy, Amplituhedronsquared}. Specifically, Marcott shows that the space parametrized by the Wilson loop diagrams is  exactly the correct positroid cell in $\Gr(k,n)$, up to a set of measure zero \cite{WLDdim}.

In the above work and elsewhere, it is clear that while the space \hlfix{parametreized by}{``filled by'' vs ``parameterized by'' in the previous paragraph} the Wilson loop diagrams is related to the Amplituhedron, they are not the same space. For instance, it is conjectured that the Amplituhedron is orientable, whereas Agarwala and Marcott show in \cite{non-orientable} that the space parametrized by the Wilson loop diagrams as a subspace of $\Gr(k, n+1)$ can be seen as a non-orientable vector bundle over some submanifold of $\Gr(k,n)$. Furthermore, the cells associated to the Wilson loop diagrams are different from the cells associated to BCFW diagrams, which can be seen by comparing the results of \cite{AmplituhedronDecomposition} and \cite{casestudy}. This is an intriguing observation, as it is conjectured that the space parametrized by the Wilson loop diagrams will project down to something close to the Amplituhedron \cite{wilsonloop, Amplituhedronsquared}.

@@

This paper continues this body of work by establishing concrete connections between the combinatorics of the Wilson loop diagram as determined by the physics, and the combinatorics of the associated positroid cell as identified in \cite{wilsonloop}. In this paper, we give an algorithm to read the Grassmann necklace, one of the standard combinatorial characterizations of the positroid cell, directly off the Wilson loop diagram. We also show that the positroid cells associated to Wilson loop diagrams are all $3k$-dimensional, where $k$ is the number of propagators in the diagram; by contrast, the BCFW cells are all $4k$-dimensional \cite{Arkani-Hamed:2013jha}. In \cite{non-orientable} the authors note that the Wilson loop diagrams only correspond to positroid cells when the gauge vector is ignored. Including the gauge vector defines a $4k$ dimensional subspace of $\mathbb{G}r_\R(k,n+1)$.\st{This apparent contradiction}\todo{I think we should say more in view of reviewer Y's comment 3 and we can probably phrase it more positively too} \st{is resolved by the non-orientability result of {\cite{non-orientable}}.}

@@

Finally, in Section~\ref{sec poles} we characterize the denominator of the integrand of a Wilson loop diagram in terms of its associated matrix and Grassmann necklace. We show that given the \hlfix{physically correct}{This was here before but can you remind me what it is hinting at?}\sanote{This  is awkward. I think I meant that there are many representations of the positroid cell; wwe want thhe representattion that is  good for the physics.} representation of the positroid cell associated to a Wilson loop diagram, the irreducible factors of the denominator are exactly the set of irreducible factors of the minors of said representation given by the Grassmann necklace. This characterization is given explicitly by Algorithm~\ref{alg WLD to denom via GN} and Theorem~\ref{thm denom}.

@@

All three of these conditions arise from the physical interpretation of Wilson loop diagrams.  The first condition comes from the fact that these diagrams represent $N^k$MHV interctions, where $k +4 <n$, see for instance the explanation in the introduction of \cite{adamo:2011pr}.  The second condition can be derived from the integrals associated to the diagrams, as in \cite[Theorem 1.13]{wilsonloop}. Note that the second condition prohibits propagators that connect two adjacent edges. In our definition, $\cP$ is a set so we may not have two propagators starting and ending on the same pair of edges, but we have lost nothing of physical value by not allowing multiple propagators ending on the same pair of edges since these would also fail the second condition. The third condition comes from the fact that these Wilson loop diagrams are a representation of planar Feynman diagrams of $N=4$ SYM theory. For a description of the relationship between planar Feynman diagram and planar Wilson loops, see the discussion in \cite[section 3]{LipsteinMason}.
\hlfix{}{I changed the previous 2 paragraphs quite a bit, to separate the math intuition from the physical reasoning.}  \sanote{fixed cittation issues. Like the  rewrite}}

@@

One may also interpret the functional $I(W)$ as associating a volume to each cell $\Sigma(W)$ \cite{wilsonloop, Amplituhedronsquared, HeslopStewart}, where the volume is determined by the external particles involved. Specifically, the integrand is written in terms of the elements of $C(W)$ as  \begin{equation} \label{eq:omega(W) def}\Omega(W) = \frac{\prod_{r=1}^{|\cP|} \prod_{v \in V(p_r)} \textrm{d}x_{p_r}}{R(W)} \;, \end{equation} multiplied by a $\delta$ function that evaluates the matrix $C(W)$ on the external particles. The denominator $R(W)$ is given in Definition~\ref{def R(W)} below. For the purposes of this paper, we restrict our attention to $\Omega(W)$ only, and ignore the $\delta$ function in the integrand of $I(W)$. The interested reader is referred to \cite{Adamo:2012xe,HeslopStewart,LipsteinMason} for more information about the integrand of $I(W)$. \sanote{not putting sections here, but I'd like a second thought -- what is this in response to?} \sanote{I'm not sure  at this point. }

For a fixed number of propagators and particles, the tree level amplitude of the Wilson loop diagrams is computed as the sum of the $I(W)$ for all $W$ of the correct size. By {\cite{Adamo:2011pv,Britto:2005fq,Arkani-Hamed:2013jha}} this sum is finite for all sets of external particles, i.e.
\bas \mathcal{A}_{k, n}^{tree} = \sum_{\substack{W = (\cP, [n]), \\ |\cP| = k}} I(W) <\infty. \eas

This would certainly hold if  the singularities in the integrands $\Omega(W)$ (the spurious poles of the theory) cancel out in the sum; in particular, if all poles arising from $R(W)$ must cancel. Specifically, it is conjectured that these poles cancel on the boundaries of the positroid cells $\Sigma(W)$, analogous to the analysis in \cite{Arkani-Hamed:2013jha}. This has been checked for particular classes of examples (e.g. \cite{casestudy, Amplituhedronsquared, HeslopStewart}) but has not been proven in general.\hlfix{}{Is this updated for the spurious vs physical poles?}  \sanote{these arre all spurioous poles. Needs  more?}

@@

Note that the positroid cells associated to the Amplituhedron are of dimension $4|\cP|$. In \cite[section 2]{non-orientable}, Agarwala and Marcott associate a Deodhar component\footnote{Since Deodhar components play no role in this paper, the interested reader is referred to \cite{non-orientable} for more details on this.} to each Wilson loop diagram, and show that the dimension of this Deodhar component is the sum of $|\cP|$ and the dimension of the associated positroid cell.
%\[|\cP| + \text{ the dimension of associated positroid cell}.\]
Combining this with Theorem~\ref{thm dim}, we see that the Deodhar component associated to a Wilson loop diagram is $4|\cP|$-dimensional, showing that the geometry underlying the Wilson loop diagrams is consistent with the geometry underlying the Amplituhedron. It is worth noting that the Amplituhedon is defined by a projection of a set of $4|\cP|$ dimensional positroid cells, while the $4|\cP|$-dimensional cells of the Wilson loop diagram lie in $\mathbb{G}r_\R$. To  date, there is no evidence to suggest that the $3|\cP|$ dimensional positive cells of $\Gr(|\cP|, n)$ defining the Wilson loop diagrams are related to the $4|\cP|$ dimensional positive cells defining the Amplituhedron.  \sanote{This para adresses point 3 of report Y. Should there be a similar paragraph in the intro? Should it be moved?} \sanote{futzed some  more}

Marcott in \cite[Theorem 8.4]{WLDdim} also gives a proof of the $3|\cP|$-dimensionality of $M(W)$ which is much more geometric. However, it is not easy to track the effect of a particular propagator in that proof. By contrast, our approach is\todo{The transition between these paragraphs has been messed up a bit} much more constructive, and elucidates how each additional propagator contributes to tthe dimension. \sannote{fiddled with this transition} By combining Algorithms~\ref{alg:put GN on WLD} and \ref{alg:GN to Le} (converting from Wilson loop diagram to Grassmann necklace, and Grassmann necklace to Le diagram respectively) we explicitly describe the effect of adding another propagator to a Wilson loop diagram in terms of the plusses of the associated Le diagrams. Recall that the dimension of a positroid cell is equal to the number of plusses in its associated Le diagram \cite[Theorem 6.5]{Postnikov}.  Thus understanding the effect of adding another propagator is sufficient to give a recursive proof of the $3|\cP|$-dimensionality of the cells.

@@

Since SYM $N=4$ is a finite theory {\cite{Alday:2007hr, ParkeTaylorids}}, the scattering amplitudes (and thus the holomorphic Wilson loops $\,\mathfrak{W}_{k,n}$) are finite. That is, for a fixed set of external particles $\cZ$ we must have
\bas \sum_{W \text{ s.t. } |\cP| = k} I(W)(\cZ) <\infty \;.\eas
This holds if all poles of the integrands cancel. As a step towards this, in this section we prove some results about the structure of the denominator $R(W)$ of the integrand, as defined in Definition~\ref{def R(W)}.
